---
title: "Infectious Disease Dynamics: Introduction to Data Analytics and Mathematical Modelling"
format: html
execute:
  echo: false
---

```{r}
#| label: session-name-editor
#| include: false
session_entries <- c(
	"Welcome and introductions|This opening session introduces participants, faculty, and the overall structure of the short course. We also provide a practical orientation to Excel and R workflows that will be used throughout the hands-on components.",
	"Outbreak response I|We cover key delay distributions used in outbreak analytics, including reporting delays, incubation periods, and serial intervals. The session focuses on how these quantities shape interpretation of surveillance data in real time.",
	"Outbreak response II|We introduce nowcasting, time-varying reproduction numbers, and short-term forecasting for operational decision support. Participants will learn how these methods connect to actionable outbreak response planning.",
	"Practical|Participants implement renewal-equation-based outbreak models and work with tools such as EpiEstim. The practical emphasizes translating incidence data into interpretable transmission metrics.",
	"Serological models I|This session introduces population-level serological modeling, with a focus on catalytic models. We discuss how seroprevalence by age can be used to infer transmission intensity over time.",
	"Practical|Participants estimate age-dependent force of infection from serological data. The exercise emphasizes model assumptions, parameter interpretation, and quality checks.",
	"Serological models II|We move to individual-level serological models to infer infection histories from longitudinal data. The session highlights how repeated measurements improve inference on immune dynamics.",
	"Practical|Participants infer IgG dynamics for RSV using data from The Gambia. The practical focuses on linking antibody trajectories to infection and exposure processes.",
	"SIR and SIS models|This session introduces the foundations of SIR and SIS models, including the meaning of R0 and Re. We build intuition for threshold behavior and epidemic control in simple compartmental systems.",
	"Practical|Participants explore a simple transmission model to test R0-based threshold concepts. The practical reinforces how parameter changes alter epidemic trajectories.",
	"Model choice|We compare deterministic and stochastic models, difference and differential equations, dynamic and static frameworks, and individual-based models. The session discusses when each model class is most appropriate for policy-relevant questions.",
	"Practical|Participants run and compare stochastic model variants to understand variability and uncertainty. The exercise emphasizes interpretation beyond single deterministic trajectories.",
	"Heterogeneity I: age|We examine age-structured transmission using contact matrices, including POLYMOD-style data. The session shows how age mixing patterns influence reproduction numbers and intervention impact.",
	"Practical|Participants identify which age groups drive transmission under age-dependent mixing assumptions. The practical links contact structure directly to R0 and targeted control strategies.",
	"Heterogeneity II: space|This session introduces spatial heterogeneity through gravity and commuter-based models, including examples with mobile phone mobility data. We discuss how movement and connectivity reshape transmission risk across locations.",
	"Practical| tbd",
	"Fitting a model|We introduce objective functions and metrics for model calibration, with emphasis on likelihood-based inference. The session explains why likelihood is useful and how optimization is used to estimate parameters.",
	"Practical|Participants compute likelihoods and fit a simple transmission model to data. The practical focuses on implementation details and diagnostics of fit quality.",
	"Identifying a fit model|We review model fit diagnostics, including trace plots and Rhat, and signs that inference needs adjustment. The session also contrasts overfitting and underfitting in applied infectious disease modeling.",
	"Practical|Participants work with intentionally poorly fitting models and iteratively improve them. The exercise develops intuition for practical tuning and troubleshooting.",
	"Vaccine models|We compare all-or-nothing and leaky vaccine formulations and their implications for population protection. The session highlights direct and indirect effects and how these alter impact projections.",
	"Practical|Participants simulate a flu vaccination model to quantify indirect protection effects. The practical links vaccine assumptions to changes in incidence and policy conclusions.",
    "Health Economics|We introduce core health economic metrics including QALY, DALY, and ICER. The session connects epidemiological model outputs to economic decision-making.",
    "Practical|Participants calculate DALYs and run a probabilistic sensitivity analysis. The exercise demonstrates uncertainty-aware evaluation of intervention value.",
	"Identifying a good model|We use the WHO IVIR framework to assess model quality, transparency, and policy relevance. The session provides a structured approach to judging whether a model is fit for decision support.",
	"Group work I|Participants form groups, define topics, and start scoping their modeling question. The session sets expectations for outputs and group workflow.",
	"Heterogeneity III: vectors|This session focuses on vector-borne disease modeling with examples from malaria, dengue, and tick-borne infections. We discuss how vector ecology and host exposure shape transmission dynamics.",
	"Practical| tbd",
	"Climate sensitive models| tbd",
	"Practical| tbd",
	"Accesibility of models|We review the Lancet Commission perspective and reporting guidelines for transparent and reproducible modeling. The session emphasizes communication standards that improve trust and uptake.",
	"Group work II|Groups discuss key insights from assigned readings and refine their project direction. The session supports peer feedback and stronger framing of policy questions.",
    "Phylogenetic inference|We introduce the principles of phylogenetic inference and how sequence data can inform transmission modeling. The session highlights where phylogenetics adds value and where limits remain.",
	"Practical|Participants work through a simple phylogenetic modeling example. The exercise focuses on interpretation of outputs for epidemiological use.",
	"AI and the future of models| tbd",
	"Group work III|Groups prepare policy-facing presentations for the following day. The session focuses on clear argumentation, assumptions, and uncertainty communication.",
	"Group presentations I|Groups present their work in a NITAG-style discussion format. Feedback centers on technical quality, relevance, and clarity for decision-makers.",
	"Group presentations II|Additional groups present and discuss findings in a NITAG-style format. The session emphasizes constructive critique and synthesis across projects.",
	"Use and abuse of models|This closing lecture reflects on responsible and irresponsible uses of models in public health. We discuss common pitfalls, misuse in communication, and practical safeguards.",
	"End of course|The final session closes the course, awards certificates, and summarizes key take-home messages. Participants reflect on next steps for applying methods in their own work."
)

entry_parts <- strsplit(session_entries, "\\|", fixed = FALSE)
session_table <- data.frame(
	session_id = paste0("session", seq_along(entry_parts)),
	session_name = vapply(entry_parts, function(x) x[1], character(1)),
	session_content = vapply(
		entry_parts,
		function(x) if (length(x) >= 2) paste(x[-1], collapse = "|") else "",
		character(1)
	),
	stringsAsFactors = FALSE
)
```

# Timetable

<style>
.week-one table {
	border-left: 6px solid #1d4e89;
	background: #f3f8ff;
	table-layout: fixed;
	width: 100%;
	border-collapse: separate;
	border-spacing: 0.65rem;
	margin: 1.1rem 0 2rem 0;
}

.week-two table {
	border-left: 6px solid #8a5a00;
	background: #fff8e8;
	table-layout: fixed;
	width: 100%;
	border-collapse: separate;
	border-spacing: 0.65rem;
	margin: 1.1rem 0 2rem 0;
}

.week-one th,
.week-two th {
	padding: 0.75rem;
	border-radius: 10px;
	background: rgba(255, 255, 255, 0.75);
	color: #000000;
}

.week-one td,
.week-two td {
	padding: 0.55rem;
	border-radius: 12px;
	background: rgba(255, 255, 255, 0.7);
	color: #000000;
}

.week-one th:first-child,
.week-two th:first-child,
.week-one td:first-child,
.week-two td:first-child {
	width: 14%;
}

.week-one th:nth-child(n+2),
.week-two th:nth-child(n+2),
.week-one td:nth-child(n+2),
.week-two td:nth-child(n+2) {
	width: 21.5%;
}

.week-one td:nth-child(n+2),
.week-two td:nth-child(n+2) {
	height: 6.5rem;
	vertical-align: top;
}

.week-one td:nth-child(n+2) a,
.week-two td:nth-child(n+2) a {
	display: block;
	min-height: 5.2rem;
	line-height: 1.2;
	word-break: break-word;
	padding: 0.8rem;
	border-radius: 10px;
	background: rgba(255, 255, 255, 0.9);
	transition: transform 0.18s ease, box-shadow 0.18s ease, background-color 0.18s ease;
	text-decoration: none;
	color: #000000;
}

.week-one td:nth-child(n+2) a.practical-tile,
.week-two td:nth-child(n+2) a.practical-tile {
	background: #e8f7ed;
}

.week-one td:nth-child(n+2) a.groupwork-tile,
.week-two td:nth-child(n+2) a.groupwork-tile {
	background: #fff1d8;
}

.week-one td:nth-child(n+2) a.firstlast-tile,
.week-two td:nth-child(n+2) a.firstlast-tile {
	background: #e6f3ff;
}

.week-one td:nth-child(n+2) a:hover,
.week-one td:nth-child(n+2) a:focus,
.week-two td:nth-child(n+2) a:hover,
.week-two td:nth-child(n+2) a:focus {
	transform: translateY(-2px);
	box-shadow: 0 6px 14px rgba(0, 0, 0, 0.12);
	background: #ffffff;
}

.week-gap {
	height: 2.25rem;
}
</style>

```{r}
#| label: timetable-builder
#| results: asis
if (nrow(session_table) != 40) {
	stop("session_table must contain exactly 40 rows.")
}

date_labels <- c(
	"19 April", "20 April", "21 April", "22 April", "23 April",
	"26 April", "27 April", "28 April", "29 April", "30 April"
)

session_ids <- session_table$session_id
session_names <- session_table$session_name
session_classes <- ifelse(
	grepl("^Practical", session_names, ignore.case = TRUE),
	"practical-tile",
	ifelse(grepl("^Group work", session_names, ignore.case = TRUE), "groupwork-tile", "")
)
session_classes[session_ids %in% c("session1", "session40")] <- trimws(
	paste(session_classes[session_ids %in% c("session1", "session40")], "firstlast-tile")
)
session_links <- sprintf(
	"<a class='session-tile %s' href='#%s'>%s</a>",
	session_classes,
	session_ids,
	session_names
)
session_grid <- matrix(session_links, nrow = 10, ncol = 4, byrow = TRUE)

render_week <- function(heading, row_ids, css_class) {
	cat("## ", heading, "\n\n", sep = "")
	cat("::: {.", css_class, "}\n", sep = "")
	cat("| Date | 09:00-10:00 | 10:30-12:00 | 13:00-14:00 | 14:30-16:00 |\n")
	cat("|---|---|---|---|---|\n")
	for (i in row_ids) {
		cat("| ", date_labels[i], " | ", paste(session_grid[i, ], collapse = " | "), " |\n", sep = "")
	}
	cat(":::\n\n")
}

render_week("19-23 April", 1:5, "week-one")
cat("::: {.week-gap}\n:::\n\n")
render_week("26-30 April", 6:10, "week-two")
```

# Sessions

```{r}
#| label: session-headings-builder
#| results: asis
for (i in seq_len(nrow(session_table))) {
	id <- session_table$session_id[i]
	name <- session_table$session_name[i]
	content <- session_table$session_content[i]
	cat("## ", name, " {#", id, "}\n\n", sep = "")
	if (nzchar(trimws(content))) {
		cat(content, "\n\n")
	}
}
```

